---
title: "Patchwork Analysis"
output: html_notebook
author: Kaia Newman
date: 04-11-2026
---


```{r}
# install required packages
# options(repos = list(CRAN="http://cran.rstudio.com/"))
# install.packages("knitr")
# install.packages("rmarkdown")
# install.packages("multcomp")
# install.packages("likert")
# install.packages("lmerTest")
# install.packages("parameters")
# install.packages("sqldf")
# install.packages("stargazer")
# install.packages("GGally")
# install.packages("broom.mixed")
# install.packages("patchwork")
# install.packages("ggpubr")
# install.packages("merDeriv")
# install.packages("AER")
# install.packages("effectsize")
# install.packages("tidyverse")
# install.packages("ggplot2")
# install.packages("performance")
# install.packages("lme4", type = "source")
# install.packages("moments")
install.packages("simr")
```

# Load libraries
```{r}
#load libraries
library(ggplot2)
library(grid)
library(gridExtra)
library(likert)
library(RColorBrewer)
library(dplyr)
library(tidyr)
library(readr)
library(car)
library(lmerTest)
library(sqldf)
library(stargazer)
library(tidyverse)
library(GGally)
library(broom.mixed)
library(patchwork)
library(ggpubr)
library(merDeriv)
library(nnet)
library(AER)
library(corrplot)
library(effectsize)
library(rstatix)
library(performance)
library(multcomp)
library(simr)
library(nnet)
library(psych)

source("patchwork_processing.R")
nrow(timing_data)
```
```{r}
# Make full df, gaze-valid df, full df had patch, gaze-valid had patch 
full_df = read.csv("timing_correctness_data.csv")

# TODO: do a bunch of preprocessing.
gaze_valid_df = full_df[!is.na(full_df$Source.Code_fixation_count), ]
full_df_had_patch = full_df[(full_df$condition == "overfitting" | full_df$condition == "correct"), ]
gaze_valid_had_patch = full_df_had_patch[!is.na(full_df_had_patch$Source.Code_fixation_count), ]
```


# RQ1a: Just run the models from APR, What is it Good For? (ICSE 2024)
```{r}
# Note: these are not our final models for this RQ, and were not derived in a principled way. We're just seeing how results generalize.
str(timing_data[, c("correct", "PID", "bug", "condition")])

correctness_model = glmer(correct ~ condition + (1 |PID) + (1 | bug), data = timing_data, family = binomial)

summary(correctness_model)

timing_model = lmer(time_minutes ~ condition + num_prev + (1 | PID) + (1 | bug), data = timing_data)

summary(timing_model)

# Redo significance testing 
k1 <- glht(timing_model, mcp(condition="Tukey"))$linfct
k2 <- glht(timing_model, mcp(num_prev="Tukey"))$linfct
posthoc <- glht(timing_model, linfct = rbind(k1,k2))
summary(posthoc)
model_parameters(timing_model)

r2(timing_model)
Anova(timing_model)
```
# RQ1a: How do patch suggestions affect time and accuracy of debugging sessions? (Section FIXME) 
```{r}
# And now, as determined by the DAG, participant insights, and prior literature, we define the maximal model
# FIXED EFFECTS:
# key players: treatment condition, professional YOE. these are the ones we'll try nested models for in LRT.
# intellij/java experience
# self-efficacy scores
# number of previous tasks, looking for ordering effects 

# Note: my intuition is that professional YOE/professional identity are too related as constructs for what we're trying to measure (latent expertise) to include both
# in the same model

# We would include treatment condition * professional YOE (hypothesis from participants: professionals might be less affected by wrong patches) and some other interaction effects, but we don't really have enough samples to be able to do this

# RANDOM EFFECTS:
# bug, PID
# optimizer maybe helps the convergence some because it goes for more iterations

full_correctness_model = glmer(correct ~ condition + professional_YOE
                              + java_intellij_experience + self_efficacy + num_prev
                              + (1 | PID) + (1 | bug), 
                              data = timing_data, 
                              family = binomial)

# Do LRT to see whether condition, professional_YOE, and professional_YOE:condition improve the model
correctness_model_no_condition = glmer(correct ~ professional_YOE
                              + java_intellij_experience + self_efficacy + num_prev
                              + (1 | PID) + (1 | bug), 
                              data = timing_data, 
                              family = binomial)
condition_LRT = anova(full_correctness_model, correctness_model_no_condition)

correctness_model_no_YOE = glmer(correct ~ condition
                              + java_intellij_experience + self_efficacy + num_prev
                              + (1 | PID) + (1 | bug), 
                              data = timing_data, 
                              family = binomial)
YOE_LRT = anova(full_correctness_model, correctness_model_no_YOE)

# BH-correct for multiple comparisons in LRTs and present results
p_vals = c(
  condition    = condition_LRT$`Pr(>Chisq)`[2],
  YOE          = YOE_LRT$`Pr(>Chisq)`[2]
)

p_adjusted = p.adjust(p_vals, method = "BH")

# print both together for easy comparison
data.frame(
  predictor = names(p_vals),
  p_raw     = p_vals,
  p_adj     = p_adjusted
)

# LRT results: YOE does not improve the model
# Summarize the best model to report
summary(correctness_model_no_YOE)

# Overall results:
# Correct patch suggestions predict solution correctness
# No evidence for overfitting patches making things worse

# Summary stats
summary(timing_data$time_minutes)

# Look at distribution shape
# FIXME: log transform? might need to use tobit?
# we worry about skewness in residuals, not necessarily the skewness of the response times untransformed. but the shape of the data should remain the same if transformed
hist(timing_data$time_minutes, breaks = 30)

full_timing_model = glmer(time_minutes ~ condition + professional_YOE
                              + java_intellij_experience + self_efficacy + num_prev
                              + (1 | PID) + (1 | bug), 
                              data = timing_data, 
                              family = Gamma(link="log"))

# Check residuals because of skewness 
residuals = resid(full_timing_model)
hist(residuals)
paste0("This is the skewness of the residuals: ", skewness(residuals))

# correcting for multiple comparisons for categorical predictors
k1 <- glht(timing_model, mcp(condition="Tukey"))$linfct
k2 <- glht(timing_model, mcp(num_prev="Tukey"))$linfct
posthoc <- glht(timing_model, linfct = rbind(k1,k2))
summary(posthoc)
model_parameters(timing_model)

r2(timing_model)
Anova(timing_model)
```
# RQ4: How do patch suggestions influence cognitive load during debugging? (Section FIXME)
```{r}
# High level plan: do subjective reports from NASA-TLX, and then see if eye tracking metrics agree with the self-reports

# do model with self-reported predictors here
cognitive_load_model = lmer(cognitive_load ~ condition + professional_YOE
                              + java_intellij_experience + num_prev
                              + (1 | PID) + (1 | bug), 
                              data = timing_data)

summary(cognitive_load_model)

# Redo significance testing
k1 <- glht(cognitive_load_model, mcp(condition="Tukey"))$linfct
k2 <- glht(cognitive_load_model, mcp(num_prev="Tukey"))$linfct
posthoc <- glht(cognitive_load_model, linfct = rbind(k1,k2))
summary(posthoc)
model_parameters(cognitive_load_model)
standardize_parameters(cognitive_load_model)

r2(cognitive_load_model)
Anova(cognitive_load_model)

# Gaze metrics related to cognitive load: average fixation duration
# related metric:
# avg_fixation_duration
fix_duration_model = glmer(avg_fixation_duration ~ had_patch + (1|PID) + (1|bug), 
                          data = timing_data,
                          family = Gamma(link="log"))
summary(fix_duration_model)
anova(fix_duration_model)
```

# Confusion matrix for patch judgments
```{r}
confusion_data <- had_patch %>%
  group_by(think_patch_correct, patch_correct) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(
    pct = round(n / sum(n) * 100, 1),
    think_label = ifelse(think_patch_correct == 1, "Think: Correct", "Think: Incorrect"),
    actual_label = ifelse(patch_correct == 1, "Actually: Correct", "Actually: Incorrect")
  )

had_patch

ggplot(confusion_data, aes(x = actual_label, y = think_label, fill = n)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(n, "\n(", pct, "%)")), size = 6) +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(x = "Actual", y = "Think", fill = "Count", title = "Confusion Matrix for Patch Judgments") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 13),
    axis.title = element_text(size = 14)
  )

# Chi^2 test between think patch correct and actually correct
patch_table = table(had_patch$condition, had_patch$think_patch_correct)
patch_table = patch_table[rownames(patch_table) != "control",]

chisq.test(patch_table)

# People are statistically significantly inclined to agree with reality

# TODO: test to see if IDK response

ggsave("patch_confusion_matrix.png", width = 6, height = 5, dpi = 300)
```

# Confusion matrix for correctness 
```{r}
confusion_data2 <- correct_no_idk %>%
  group_by(think_correct, correct) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(
    pct = round(n / sum(n) * 100, 1),
    think_label = ifelse(think_correct == 1, "Think: Correct", "Think: Incorrect"),
    actual_label = ifelse(correct == 1, "Actually: Correct", "Actually: Incorrect")
  )

ggplot(confusion_data2, aes(x = actual_label, y = think_label, fill = n)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(n, "\n(", pct, "%)")), size = 6) +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(x = "Actual", y = "Think", fill = "Count", title = "Confusion Matrix for Solution Judgments") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 13),
    axis.title = element_text(size = 14)
  )

# Chi^2 test between actual correctness and think their solution is correct
correct_table = table(correct_no_idk$correct, correct_no_idk$think_correct)

res = chisq.test(correct_table)
res
res$stdres

# Indeed, people are statistically significantly inclined to agree with reality

# Total for accuracy of correctness judgments
judgment_accuracy = 38.7 + 40.5
paste0("Total accuracy of correctness judgments: ", judgment_accuracy, "%")

ggsave("confusion_matrix.png", width = 6, height = 5, dpi = 300)

# FIXME: do analyses of whether or not professionals are more likely to say 'IDK'
```

# Testing hypotheses for RQ2: What debugging strategies and behaviors distinguish successful from unsuccessful debugging attempts? (Section FIXME)
# Hypotheses derived from interview data
```{r}
# TODO: figure out survivor analysis with anything where the outcome variable is time

# FIXME: broken, skewness won't work and might be masked by a different package (parameters)
# helper function to check if model residuals are normally distributed
check_residuals <- function(model) {
  model_name <- deparse(substitute(model))
  resids     <- residuals(model)
  skew       <- skewness(resids)
  
  par(mfrow = c(1, 2))
  hist(resids, main = paste("Residuals:", model_name), xlab = "Residuals")
  qqnorm(resids, main = model_name)
  qqline(resids, col = "red")
  par(mfrow = c(1, 1))
  
  cat(model_name, "- residual skewness:", round(skew, 3), "\n")
  if (abs(skew) > 0.5) {
    cat("  ^ consider a different model family or transform response var\n")
  }
}

# Hypothesis: people who look at the patch early on may be anchored to the patch's solution and may finish the task faster/more correctly
# expect condition:TTFF interaction ~ correctness and TTFF ~ time in glmer model
# TODO: make sure it's dropping all the relevant columns

# ttff_patch, avg_fixation_duration_patch (~ correctness)
ttff_patch_model = glmer(correct ~ ttff_patch + (1|PID) + (1|bug), 
                 data = timing_data, family = binomial)
summary(ttff_patch_model)

ttff_patch_model_time = lmer(time_minutes ~ ttff_patch + (1|PID) + (1|bug), 
                 data = timing_data)
summary(ttff_patch_model_time)
anova(ttff_patch_model_time)

# Hypothesis: Some developers described spending a while evaluating the patch and developing suspicion about it. Developers who look at the patch for longer or more carefully may be more likely to debug correctly or quickly. 
# fixation %/count and average fixation duration ~ correctness
# TODO: putting condition just for correctness since it appears to differ between overfitting/correct?
avg_dur_patch_model = glmer(correct ~ avg_fixation_duration_patch + condition + (1|PID) + (1|bug), 
                 data = timing_data, family = binomial)
summary(avg_dur_patch_model)

avg_dur_patch_model_time = lmer(time_minutes ~ avg_fixation_duration_patch + (1|PID) + (1|bug), 
                 data = timing_data)
summary(avg_dur_patch_model_time)
anova(avg_dur_patch_model_time)

# Hypothesis: Developers report triangulating multiple sources of information (tests, test output, code, patch, web-search) to debug, and prior work supports the idea that higher attention switching behavior correlates with accurate proof correction/problem-solving. Developers who exhibit higher attention switching behavior between AOIs may be more likely to debug correctly.
# look at number of cross-AOI switches / time ~ correctness
attention_switch_model = glmer(correct ~ attention_switching_rate + (1|PID) + (1|bug), 
                 data = timing_data, family = binomial)
summary(attention_switch_model)

attention_switch_model_time = lmer(time_minutes ~ attention_switching_rate + (1|PID) + (1|bug), 
                 data = timing_data)

# Hypothesis: Overwhelmingly, developers report having access to a patch speeding up fault localization. When developers have access to a patch, do they find the location of the bug faster and search more efficiently? 
# TTFF on buggy method ~ has_patch
# fixation rate ~ has_patch 
# check for subset of trials on Chart 12 -- this one had an overfitting patch that was not in the correct method
# First do logistic regression: did having a patch help people look at the buggy method more at all? 
# Then, try a survival analysis -- we have good reason to try this since we're trying to predict the time until an event occurs, with right-censored data
# TODO: ask Marie about fixation_rate: the intent was to see if having a patch made them have more efficient search, yes? is this for cognitive load RQ? 
fl_model = glmer(looked_at_buggy_method ~ had_patch + (1|PID) + (1|bug), 
                 data = timing_data, family = binomial)
summary(fl_model)

fix_rate_model = lmer(fixation_rate ~ had_patch + (1|PID) + (1|bug), data = timing_data)
summary(fix_rate_model)
anova(fix_rate_model)

# unfortunately, don't have enough data to just try this for chart12 since it embedded incorrect FL info

# Hypothesis: Along similar lines, developers also report having access to a patch narrowing their focus. When developers have access to a patch, do they look at a more narrow slice of a codebase or have a "pointier" distribution of attention?
# num_files_looked_at (should account for time confound), fixation_duration_entropy
# an offset just makes the coefficient 1, so creates assumption where there is a 1:1 relationship between the amount of time spent on task and the number of files you look at
num_files_m <- glmer(num_files_looked_at ~ had_patch + offset(log(time_minutes)) + (1|PID) + (1|bug), 
           data = timing_data, family = poisson)
summary(num_files_m)

entropy_m = lmer(fixation_duration_entropy ~ had_patch + (1|PID) + (1|bug), 
                 data = timing_data)
summary(entropy_m)

# TODO check all the residuals like so:
# check_residuals(entropy_m)
```


# ART ANOVA for:
# RQ2: What debugging strategies and behaviors distinguish successful from unsuccessful debugging attempts? (Section FIXME)
# RQ3: How do patch suggestions influence developers' debugging strategies and behaviors? (Section FIXME)
```{r}
library("ARTool")
og_timing_data = timing_data
# TODO: determine whether '-' and 'Unknown/OOB' AOIs are MNAR by drawing pic. comment on this in paper

# Then, remove the "Patch" AOI for gaze distributions between correct vs. incorrect trials
# This is appropriate given our RQ targets how debugging behaviors /change/ in response to a patch, so we expect there to be some difference on the distribution
# over other AOIs

# TODO: also look at differences between when people got the problem correct/wrong just for when they had a patch 

# Find count and duration columns
fixation_count_cols    <- grep("fixation_count", names(timing_data), value = TRUE)
fixation_duration_cols <- grep("fixation_duration", names(timing_data), value = TRUE)

# Exclude other duration metrics from columns
# And also exclude patch fixations from the columns, since we're only interested in seeing how this affected other stages of debugging
fixation_duration_cols = fixation_duration_cols[fixation_duration_cols != "avg_fixation_duration"
                                                 & fixation_duration_cols != "avg_fixation_duration_patch"
                                                 & fixation_duration_cols != "fixation_duration_entropy"
                                                 & fixation_duration_cols != "Patch_fixation_duration"
                                                 & fixation_duration_cols != "OOB_fixation_duration"
                                                 & fixation_duration_cols != "X._fixation_duration"]

# Exclude missing data for AOIs or fixations outside of AOIs we consider
fixation_count_cols = fixation_count_cols[fixation_count_cols != "Patch_fixation_count"
                                          & fixation_count_cols != "X._fixation_count"
                                          & fixation_count_cols != "OOB_fixation_count"]

# fill 0s for NAs left in the fixation_counts/durations columns per row because ART can't work on NAs
timing_data[fixation_count_cols][is.na(timing_data[fixation_count_cols])] = 0
timing_data[fixation_duration_cols][is.na(timing_data[fixation_duration_cols])] = 0

# Create columns for the total counts and durations, then columns for the proportions of counts and durations per AOI
timing_data$total_fixation_count    <- rowSums(timing_data[, fixation_count_cols], na.rm = TRUE)
timing_data$total_fixation_duration <- rowSums(timing_data[, fixation_duration_cols], na.rm = TRUE)
timing_data[, paste0("prop_", fixation_count_cols)]    <- timing_data[, fixation_count_cols] / timing_data$total_fixation_count
timing_data[, paste0("prop_", fixation_duration_cols)] <- timing_data[, fixation_duration_cols] / timing_data$total_fixation_duration

# Subset to only participant-tasks where we had valid gaze data
timing_data <- timing_data[timing_data$total_fixation_count > 0, ]
print(nrow(timing_data))
table(timing_data$correct)
table(timing_data$had_patch)

run_art_analysis <- function(timing_data, prop_cols, group_var, metric) {

  # makes 'AOI' a column so there's a row for each one per participant-task
  timing_data_long <- reshape(
    timing_data[, c("PID", "bug", group_var, prop_cols)],
    varying   = prop_cols,
    v.names   = "proportion",
    timevar   = "AOI",
    times     = prop_cols,
    direction = "long"
  )
  
  rownames(timing_data_long) <- NULL
  
  # remove leading crap from AOI names
  timing_data_long$AOI <- gsub("prop_|_fixation_count|_fixation_duration", "", timing_data_long$AOI)

  timing_data_long$AOI            <- factor(timing_data_long$AOI)
  timing_data_long[[group_var]]   <- factor(timing_data_long[[group_var]])
  timing_data_long$PID <- factor(timing_data_long$PID)
  
# Visualization code 
# Compute mean proportion per AOI per group
means <- aggregate(proportion ~ AOI + get(group_var), data = timing_data_long, FUN = mean)
names(means)[2] <- group_var

# Reshape to wide for barplot
means_wide <- reshape(means,
  idvar     = "AOI",
  timevar   = group_var,
  v.names   = "proportion",
  direction = "wide"
)
rownames(means_wide) <- means_wide$AOI
means_wide$AOI <- NULL
# \\. is an escape sequence for . so it means a literal dot instead of a regex
colnames(means_wide) <- gsub("proportion\\.", "", colnames(means_wide))

# Plot
colors <- rainbow(nrow(means_wide))
barplot(as.matrix(means_wide),
  col     = colors,
  legend  = rownames(means_wide),
  beside  = FALSE,
  main    = paste0("Proportion of fixation ", metric, "s by AOI and ", group_var),
  xlab    = group_var,
  ylab    = "Mean proportion",
  args.legend = list(x = "topright")
)
  # Run the analysis
  formula <- as.formula(paste("proportion ~", group_var, "* AOI + (1|PID) + (1|bug)"))
  m <- art(formula, data = timing_data_long)
  return(list(model = m, anova = anova(m), data = timing_data_long))
}

# Seeing the difference in proportions between fixation counts and durations over AOIs given patch vs. no patch, correct vs. incorrect
count_prop_cols    <- paste0("prop_", fixation_count_cols)
duration_prop_cols <- paste0("prop_", fixation_duration_cols)

res1 <- run_art_analysis(timing_data, count_prop_cols,    "correct", "count")
res2 <- run_art_analysis(timing_data, count_prop_cols,    "had_patch", "count")
res3 <- run_art_analysis(timing_data, duration_prop_cols, "correct", "duration")
res4 <- run_art_analysis(timing_data, duration_prop_cols, "had_patch", "duration")


# Correct for multiple comparisons across interaction p-values
extract_pvalues <- function(res, analysis_name) {
  if (is.null(res$anova)) return(NULL)
  
  pvals <- res$anova$`Pr(>F)`
  effects <- rownames(res$anova)
  
  data.frame(
    analysis = analysis_name,
    effect   = effects,
    p_value  = pvals
  )
}

# Collect all p-values
pval_df <- rbind(
  extract_pvalues(res1, "count ~ correct"),
  extract_pvalues(res2, "count ~ had_patch"),
  extract_pvalues(res3, "duration ~ correct"),
  extract_pvalues(res4, "duration ~ had_patch")
)

# subset to interaction effects since those are our RQs and adjust
interaction_rows <- grepl(":", pval_df$effect)
pval_df$p_adjusted <- NA
pval_df$p_adjusted[interaction_rows] <- p.adjust(pval_df$p_value[interaction_rows], method = "BH")

# Print
print(pval_df)
```

